---
title: "Memory-bounded fire-regime summaries with fireregimetools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Memory-bounded fire-regime summaries with fireregimetools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4)
```

```{r setup}
library(fireregimetools)
```

## The workflow

A fire-regime analysis runs the same landscape many times (replicates) and
characterises how fire behaves across them. With large study areas and many
replicates the per-event fire tables do not fit comfortably in memory all at
once, so fireregimetools keeps the reduction memory-bounded:

1. write each replicate's fire table to its own partition of an on-disk Arrow
   dataset (`write_burn_parquet()`);
2. open the dataset lazily (`open_burn_dataset()`) and reduce across replicates by
   pushing the aggregation down to Arrow compute
   (`summarize_fire_sizes()`, `summarize_annual_area()`, `fire_cycle()`), so the
   per-record rows are never all held in memory at once;
3. plot the fire-size distribution (`fire_size_histogram()`).

## Fire sizes

Write a few replicates' per-event fire-size tables, then summarise the
across-replicate distribution:

```{r fire-sizes}
set.seed(1)
root <- file.path(tempdir(), "fire-sizes")
unlink(root, recursive = TRUE)

for (rep_id in 1:5) {
  events <- data.frame(
    scenario = "hrv",
    year = sample(1:100, 40, replace = TRUE),
    size = rlnorm(40, meanlog = 3, sdlog = 1.5)
  )
  write_burn_parquet(events, root, replicate = rep_id)
}

summarize_fire_sizes(root)
```

Only the small summary is materialised in R; the per-event rows stay on disk. The
fire-size distribution is drawn from the pooled events:

```{r histogram}
events_all <- dplyr::collect(open_burn_dataset(root))
fire_size_histogram(events_all, size_col = "size")
```

## Annual area burned and fire cycle

For a per-replicate annual-area-burned table, `summarize_annual_area()` returns
the across-replicate envelope per time step and region, and `fire_cycle()`
converts mean annual area burned into a landscape fire cycle (years):

```{r annual}
root2 <- file.path(tempdir(), "annual-area")
unlink(root2, recursive = TRUE)

for (rep_id in 1:5) {
  annual <- data.frame(
    scenario = "hrv",
    time = rep(seq(0, 200, by = 100), each = 1),
    region = "all",
    area_ha = runif(3, 100, 2000)
  )
  write_burn_parquet(annual, root2, replicate = rep_id)
}

summarize_annual_area(root2)

fire_cycle(root2, landscape_ha = 1e6)
```
